# [Model 3] Linear Regression

```{r}
library(readxl)
library(tidyverse)
library(lubridate)
library(randomForest)
library(datasets)
library(caret)
library(rpart)

data <- read.csv('online_retail_cleaned_adjusted.csv')
# str(data)
```

```{r}
# train test split
set.seed(111)
ind <- sample(2, nrow(data), replace = TRUE, prob = c(0.75, 0.25))
trainset<- data[ind==1, ]
trainset <- subset(trainset, select = -X)
testset <- data[ind==2, ]
testset <- subset(testset, select = -X)
```

## Perform Linear Regression

```{r}
start_time <- Sys.time()
lr <- lm(Sales~., data = trainset)
end_time <- Sys.time()
time3 <- end_time-start_time

summary(lr)
```

## Linear Regression Evaluation

```{r}
plot(predict(lr, newdata = subset(testset, select = -Sales)), testset$Sales,
     main = 'Predicted values vs. True values')
pred <- predict(lr, newdata = subset(testset, select = -Sales))
print(paste('The RMSE is:', RMSE(pred = pred, obs = testset$Sales)))
```

We can simply interpret the the liner model by reading its summary. According to our sampled data, every unit increase in `Quantity` increases total sales by 0.65. The total sale will increase by 1.26 for every unit increase in `UnitPrice`. The total sale will reduce by 0.38 for every unit increase in `time`. However, in a linear regression model, the factors `month` and `week` have no significance in predicting sales.The p-values for `month` and `week` also suggests that there is no association between total sales and these two factors in a linear regression model.

The RMSE is 7.54, which is the highest compared to the previous two models. From the predicted values vs. true values plot, the linear model underestimate the total sales.

## Interpretation of Linear Model

```{r}
library(vip)
vip(lr, method ='firm')
```

The Variance Importance Plot of Linear Regression model is similar to Random Forest Regression model. `Quantity` and `UnitPrice` are still two most important variable in predicting `Sales`. The importance of `time` is low, but not as negligible as it was in Tree Regression model.
